### 2016-12-09 exampleandimage/egbibinfoot.tex
Add files via upload

\documentclass{article}
\usepackage{ctex}
\usepackage[paperwidth=12cm,paperheight=9cm,%
left=1cm,right=1cm,top=1cm,bottom=1.5cm]{geometry}
\usepackage[colorlinks=true,pdfstartview=FitH,linkcolor=blue,
anchorcolor=violet,citecolor=magenta]{hyperref}%书签功能，选项去掉链接红色方框
\usepackage{titleref}%标题引用
\usepackage[backend=biber,style=gb7714-2015ay]{biblatex}
\renewcommand{\bibfont}{\zihao{6}}
\addbibresource[location=local]{example.bib}

\begin{document}

\section{页脚的文献表}
参考文献\cite{白书农1998-146-163}
\footfullcite{陈志勇2011--}
\footfullcite{储大同2010-721-724}
\footfullcite{顾炎武1982--}
中文段落参考文献中文段落参考文献中文段落参考文献中文段落参考文献中文段落
参考文献中文段落参考文献中文段落参考文献中文段落参考文献中文段落参考文献
中文段落参考文献中文段落参考文献中文段落参考

\printbibliography[title=【全部引文】]
\end{document}