### 2017-01-05 exampleandimage/egbeamersn.tex
Add files via upload

\documentclass{beamer}
\usepackage{chemformula}
\usepackage[utf8]{inputenc}
\usepackage[german, english]{babel} % required for rendering German special characters
\usepackage{siunitx} %pretty measurement unit rendering
\usepackage{hyperref} %enable hyperlink for urls
\usepackage{ragged2e}
\usepackage[font=scriptsize,justification=justified]{caption}
\usepackage{array,booktabs,tabularx}
\usepackage[backend=biber,bibstyle=gb7714-2015-m,%nature,%
citestyle=gb7714-2015,%,backref=true%gb，gbt7714_2005_n，
align=gb7714-2015%
]{biblatex}%biber，caspervector
\renewcommand{\bibfont}{\footnotesize}
\addbibresource[location=local]{bib.bib}

\title{\huge Opto-fMRI of Monoaminergic Systems and Psychotropic Drugs}
\author{Horea-Ioan Ioanas$^{1,2}$, Bechara Saab$^{2}$, Markus Rudin$^{1}$}
\institute[ETH]{$^{1}$Institute for Biomedical Engineering, ETH and University of Zurich \\ $^{2}$Preclinical Laboratory for Translational Research into Affective Disorders, DPPP, Psychiatric Hospital, University of Zurich}
\date{\today}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\fbox{img/ag1}%
reference
\cite{Paivi}
\cite{Oegren2008}
\end{frame}

\begin{frame}
\begin{block}{References}
\footnotesize
\printbibliography[heading=bibliography,title=bibliography]
\end{block}
\end{frame}

\begin{frame}{References}
\footnotesize
\printbibliography[heading=bibliography,title=bibliography]
\end{frame}

\end{document}