### 2017-10-31 exampleandimage/egzazhi.tex
v1.0 d

\documentclass[twocolumn]{article}
\usepackage{ctex}
\usepackage{geometry}
\geometry{papersize={21cm,18cm},top=2.5cm, bottom=2.5cm, left=1cm,right=1cm}
\usepackage{xcolor}
\usepackage[CJKbookmarks,colorlinks,bookmarksnumbered=true,pdfstartview=FitH,linkcolor=blue]{hyperref}
\usepackage[backend=biber,style=zazhi]{biblatex}%

\addbibresource[location=local]{egbib.bib}
%\renewcommand{\bibfont}{\zihao{6}}
%\setlength{\bibitemsep}{0ex}
%\setlength{\bibnamesep}{0ex}
%\setlength{\bibinitsep}{0ex}

\begin{document}

\newpage
\section{bibliography}
\begin{refsection}
one author\cite{Harker1987-20-25}
\cite{Hsin1990-15-20,Prigogine1976-93-134}
\cite{USL2001--,,Cone1963--,ISO1982--}

two authors\cite{Carter2011--}

three or more authors\cite{Hino1993-173-193,Miller1984-435-454,Miller1984-435-454a}

\printbibliography[heading=bibliography]
\end{refsection}

\newgeometry{top=1.5cm, bottom=1.5cm, left=1cm,right=1cm}
\section{文献表}
\begin{refsection}
文献\cite{McEliece1984-44-53,Mittelbach2004--,MARTIN1996-85-96,
FOURNEY1971-17-38,HOPKINSON--,TACHIBANA2002--,Eggrers--,ISO1982--,CALMS1965--}

\printbibliography[heading=bibliography]
\end{refsection}

\end{document}