### 2018-06-04 exampleandimage/testfilterCategory.tex
update-20180604

\documentclass[twoside]{article}
\usepackage{ctex}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage[top=10pt,bottom=10pt,left=10pt,right=10pt]{geometry}
%\usepackage[backend=biber,style=gb7714-2015,gbalign=gb7714-2015,
%gbnamefmt=none,gbfieldstd=false,gbtype=true,gbpunctin=true,gbpub=true]{biblatex}

\usepackage[backend=biber,style=gb7714-2015,gbalign=center]{biblatex}
%gbnamefmt=givenahead,lowercase,familyahead,pinyin

\usepackage{filecontents}
\begin{filecontents}{\jobname.bib}
@ARTICLE{fazzari1993working,
  AUTHOR = {Fazzari, Steven M and Petersen, Bruce C},
  PUBLISHER = {JSTOR},
  DATE = {1993},
  JOURNALTITLE = {The RAND Journal of Economics},
  PAGES = {328--342},
  TITLE = {Working capital and fixed investment: new evidence on financing constraints},
  nameformat={lowercase},
}%namefmtid={1},

@INCOLLECTION{adam2011house,
  AUTHOR = {Adam, Klaus and Kuang, Pei and Marcet, Albert},
  PUBLISHER = {University of Chicago Press},
  BOOKTITLE = {NBER Macroeconomics Annual 2011, Volume 26},
  DATE = {2011},
  PAGES = {77--122},
  TITLE = {House price booms and the current account},
}%namefmtid={0},

@ARTICLE{bernanke1989agency,
  AUTHOR = {Bernanke, Ben and Gertler, Mark},
  PUBLISHER = {JSTOR},
  DATE = {1989},
  JOURNALTITLE = {The American Economic Review},
  PAGES = {14--31},
  TITLE = {Agency costs, net worth, and business fluctuations},
  nameformat={givenahead},
}%namefmtid={2},

@REPORT{calvo2006phoenix,
  AUTHOR = {Calvo, Guillermo A and Izquierdo, Alejandro and Talvi, Ernesto},
  INSTITUTION = {National Bureau of Economic Research},
  DATE = {2006},
  TITLE = {Phoenix miracles in emerging markets: recovering without credit from systemic financial crises},
  nameformat={familyahead},
}%namefmtid={3},

@ARTICLE{wang2006another,
  AUTHOR = {Wang, Peng-fei and Wen, Yi},
  PUBLISHER = {Elsevier},
  DATE = {2006},
  JOURNALTITLE = {Journal of Economic Dynamics and Control},
  NUMBER = {12},
  PAGES = {2533--2552},
  TITLE = {Another look at sticky prices and output persistence},
  VOLUME = {30},
  nameformat={pinyin},
}%namefmtid={4},

@REPORT{xu2012role,
  AUTHOR = {Xu, TengTeng},
  INSTITUTION = {Faculty of Economics, University of Cambridge},
  DATE = {2012},
  TITLE = {The role of credit in international business cycles},
  nameformat={pinyin},
}

@ARTICLE{zhang2009china,
  AUTHOR = {Zhang, Wen lang},
  PUBLISHER = {Elsevier},
  DATE = {2009},
  JOURNALTITLE = {Journal of Macroeconomics},
  NUMBER = {3},
  PAGES = {473--484},
  TITLE = {China’s monetary policy: Quantity versus price rules},
  VOLUME = {31},
  nameformat={pinyin},
}

    @Book{Yi2013--,
      Title                    = {Supersonic and hypersonic nozzle design},
      Address                  = {BeiJing},
      Author                   = {Yi, Shi-He and Zhao, Yu Xin and He, Lin and Zhang, Ming Li},
      Publisher                = {National Defense Industry Press},
      Year                     = {2013},
      nameformat={pinyin},
    }

    @Book{易仕和2013--,
  Title                    = {超声速和高超声速喷管设计},
  Address                  = {北京},
  Author                   = {易仕和 and 赵玉新 and 何霖 and 张敏莉},
  Publisher                = {国防工业出版社},
  Year                     = {2013}
}

    @Periodical{AAAS1883--,
  Title                    = {Science},
  Date                     = {1883},
  Editor                   = {{American Association for the Advancement of Science}},
  Number                   = {1},
  Volume                   = {1},
  Location                 = {Washington, D.C.},
  Publisher                = {American Association for the Advancement of Science}
}

@Proceedings{ROSENTHALL1963--,
  Title                    = {Proceedings of the Fifth Canadian Mathematical Congress, University of Montreal, 1961},
  Date                     = {1963},
  Publisher                = {University of Toronto Press},
  Editor                   = {E M ROSENTHALL},
  Location                 = {Toronto}
}

@Thesis{CALMS1965--,
  Title                    = {Infrared spectroscopic studies on solid oxygen},
  Author                   = {R B CALMS},
  Date                     = {1965},
  Institution              = {Univ. of California},
  Location                 = {Berkeley}
}

@Report{WHO1970--,
  Title                    = {Factors regulating the immune response},
  Author                   = {{World Health Organization}},
  Date                     = {1970},
  Institution              = {WHO},
  Location                 = {Geneva},
  Titleaddon               = {report of WHO Scientific Group}
}

@Inproceedings{FOURNEY1971-17-38,
  Title                    = {Advances in holographic photoelasticity},
  Author                   = {M E FOURNEY},
  Booktitle                = {Symposium on Applications of Holography in Mechanics, August 23-25, 1971, University of Southern California, Los Angeles, California},
  Date                     = {1971},
  Pages                    = {17-38},
  Publisher                = {ASME},
  Location                 = {New York}
}

@Article{Zhang2007-500-503,
  Title                    = {The design and experimental investigations of supersonic length shorted nozzle},
  Author                   = {Zhang, Min-li and Yi, Shi he and Zhao, Yuxin},
  Journal                  = {ACTA AERODYNAMICA SINICA},
  Number                   = {4},
  Pages                    = {500-503},
  Volume                   = {25},
  Year                     = {2007},
  nameformat={pinyin},
}

    @Article{张2007-500-503,
      Title                    = {超声速短化喷管的设计和试验研究},
      Author                   = {张敏莉 and 易仕和 and 赵玉新 and 李明},
      Journal                  = {空气动力学报},
      Year                     = {2007},
      Number                   = {4},
      Pages                    = {500-503},
      Volume                   = {25},
      language                 = {chinese}
    }
\end{filecontents}
\addbibresource{\jobname.bib}

\defbibentryset{bilangyi2013}{易仕和2013--,Yi2013--}
\defbibentryset{bilangzhang2007}{张2007-500-503,Zhang2007-500-503}

\DeclareBibliographyCategory{kindsofnameformats}
\addtocategory{kindsofnameformats}{zhang2009china,xu2012role,wang2006another,
calvo2006phoenix,bernanke1989agency,adam2011house,fazzari1993working}

    \begin{document}
    \section{作者姓名格式}

	\nocite{*}

    \printbibliography[category=kindsofnameformats,title=五种不同的姓名格式]

    \printbibliography[notcategory=kindsofnameformats,title=某种期刊的姓名格式]

    \end{document}