### 2018-06-22 exampleandimage/egfontgreak.tex

\documentclass{article}
\usepackage{ctex}
%\usepackage{fontspec}
%\setmainfont{CMU Serif}
\usepackage{geometry}
\geometry{paperwidth=15cm,paperheight=12cm,top=0cm, bottom=1.5cm, left=1cm,right=1cm}
\usepackage[backend=biber,style=gb7714-2015]{biblatex}%sorting=none,

\usepackage{filecontents}
\begin{filecontents}{\jobname.bib}
@ARTICLE{aritileeg,
  AUTHOR = {Zhang, Qi},
  DATE = {2007},
  JOURNALTITLE = {An periodical αβγδεζηθικλμνξοπρστυφχψω},
  PAGES = {82--92},
  TITLE = {An article for test αβγδεζηθικλμνξοπρστυφχψω},
  VOLUME = {12},
  NUMBER ={2},
}
\end{filecontents}
\addbibresource[location=local]{\jobname.bib}

\begin{document}

\nocite{*}
{
\newcommand{\ftcmu}{\fontspec{CMU Serif}\selectfont}
\renewcommand{\bibfont}{\ftcmu}%设置英文字体不影响中文字体
\printbibliography

}
\end{document}