### 2018-07-16 exampleandimage/testautopunct.tex
update-20180716

\documentclass[twoside]{article}
\usepackage{ctex}
\usepackage{geometry}
\geometry{paperwidth=21cm,paperheight=29cm,top=1.5cm, bottom=1.5cm, left=2.54cm,right=2.54cm}
\usepackage{xcolor}
\usepackage[CJKbookmarks,colorlinks,bookmarksnumbered=true,pdfstartview=FitH,linkcolor=blue]{hyperref}
%\usepackage[backend=biber,style=gb7714-2015,maxbibnames=5,minbibnames=5]{biblatex}
\usepackage[backend=biber,style=numeric,autocite=footnote,autopunct=true]{biblatex}%align=gb7714-2015
\usepackage{filecontents}
\begin{filecontents}{\jobname.bib}

@Inproceedings{refa,
  Title                    = {Interference cartography for hierarchical dynamic spectrum access},
  Address                  = {Chicago, IL},
  Author                   = {AlayaFeki, A. and B. Sayrac and S. Ben Jemaa and E. Moulines},
  editor={Name, B and Dome, C},
  Booktitle                = {Proc. IEEE Symp. on New Frontiers in Dynamic Spectrum Access Networks (DySPAN)},
  Date                     = {2008-10},
  Pages                    = {1-5}
}

@Inproceedings{refb,
  Title                    = {Interference cartography for hierarchical dynamic spectrum access},
  Address                  = {Chicago, IL},
  editor={Name, B and Dome, C},
  Author                   = {BlayaFeki, A. and B. Sayrac and S. Ben Jemaa and E. Moulines},
  Booktitle                = {Proc. IEEE Symp. on New Frontiers in Dynamic Spectrum Access Networks (DySPAN)},
  Date                     = {2008-10},
  Pages                    = {1-5}
}

@Inproceedings{refc,
  Title                    = {Interference cartography for hierarchical dynamic spectrum access},
  Address                  = {Chicago, IL},
  bookauthor={Name, B and Dome, C},
  Author                   = {ClayaFeki, A. and B. Sayrac and S. Ben Jemaa and E. Moulines},
  Booktitle                = {Proc. IEEE Symp. on New Frontiers in Dynamic Spectrum Access Networks (DySPAN)},
  Date                     = {2008-10},
  Pages                    = {1-5}
}

@Inproceedings{refd,
  Title                    = {Interference cartography for hierarchical dynamic spectrum access},
  Address                  = {Chicago, IL},
  bookauthor={Name, B and Dome, C},
  Author                   = {MlayaFeki, A. and B. Sayrac and S. Ben Jemaa and E. Moulines},
  Booktitle                = {Proc. IEEE Symp. on New Frontiers in Dynamic Spectrum Access Networks (DySPAN)},
  Date                     = {2008-10},
  Pages                    = {1-5}
}

@Book{Simon2001--,
  Title                    = {Spread Spectrum Communications Handbook},
  Address                  = {New York City},
  Author                   = {Simon, M. K. and Omura,J. K. and Scholtz,R. A. and Levitt,B. K.},
  Publisher                = {McGraw-Hill, Inc.},
  Year                     = {2001}
}
\end{filecontents}
\addbibresource{\jobname.bib}

\begin{document}
%编译信息:biblatex v\versionofbiblatex;gb7714-2015*.* version number \versionofgbtstyle.

\section{文献表}
文献\cite{refc}.

文献\autocite{refa}.

文献\autocite{refb}。

文献\cite{refd}.

文献\cite{Simon2001--}.

\printbibliography%[heading=subbibintoc,title=【参考文献】]
\end{document}