### 2018-07-16 exampleandimage/testindexofbib.tex

\documentclass[twoside]{article}
\usepackage{ctex}
\usepackage{geometry}
\geometry{paperwidth=21cm,paperheight=29cm,top=1.5cm, bottom=1.5cm, left=2.54cm,right=2.54cm}
\usepackage{xcolor}
\usepackage[CJKbookmarks,colorlinks,bookmarksnumbered=true,pdfstartview=FitH,linkcolor=blue]{hyperref}
\usepackage[backend=biber,style=gb7714-2015,gbtype=true,block=none,indexing=true]{biblatex}%align=gb7714-2015
\addbibresource[location=local]{example.bib}
\usepackage{makeidx}
\makeindex

\begin{document}

\section{文献表}
文献
\cite{张伯伟2002--}
\cite{白书农1998-146-163}
\cite{杨洪升2013-56-75}
\cite{中华医学会湖北分会1984--}
\cite{雷光春2012--}
\cite{贾东琴2011-45-52}
\cite{汤万金2013-09-30--}
\cite{韩吉人1985-90-99}
\cite{马欢2011-27-27}
\cite{张凯军2012-04-05--}
\cite{国家环境保护局科技标准司1996-2-3}
\cite{中国职工教育研究会1985--}
\cite{丁文祥2000--}
\cite{李强2012-05-03--}

\cite{OBRIEN1994--}
\cite{FOURNEY1971-17-38}
\cite{Park2010-696-715}
\cite{Babu2014--}
\cite{Calkin2011-8-9}
\cite{CALMS1965--}
\cite{KOSEKI2002--}
\cite{standardinfoiso158}
\cite{Dublin2012-06-14--}

\printbibliography
\printindex
\end{document}