### 2020-06-02 egbibtex/eg-thebibliography-num-include.tex

\documentclass[twoside]{article}
\usepackage{ctex}
\usepackage{geometry}
\geometry{paperwidth=21cm,paperheight=29cm,top=1.5cm, bottom=1.5cm, left=2.54cm,right=2.54cm}
\usepackage{xcolor}
\usepackage[CJKbookmarks,colorlinks,bookmarksnumbered=true,pdfstartview=FitH,linkcolor=blue]{hyperref}
\usepackage[numbers]{natbib}
%\bibliographystyle{unsrt}%GBT7714−2005AYLang−UTF8,acm,amsplain,siam,unsrt,ieeetr

\begin{document}

\include{eg-thebibliography-num-seca}

\include{eg-thebibliography-num-secb}

\end{document}